### 2022-04-27 outputs/eda.Rmd
updated scripts added data content to paper updated raw data dictionary

---
title: "exploratory data analysis"
author: "HC"
date: "21/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(here)
library(forcats)
library(modelsummary)
```

```{r}
cius_data <- read_csv(here("outputs/data/cleaned_data_v2.csv"))
cius_data[] <- lapply(cius_data, factor)
```

frequency plot of province
```{r}
ggplot(cius_data, aes(x=fct_infreq(province)))+
  geom_bar(stat = "count")
```
frequency plot of education
```{r}
ggplot(cius_data, aes(x=fct_infreq(education_level)))+
  geom_bar(stat = "count")
```
frequency plot of age group
```{r}
ggplot(cius_data, aes(x=fct_infreq(age_group)))+
  geom_bar(stat = "count")
```
series of graphs of gender, language prof, employment, immigrant status
```{r}
par(mfrow=c(2,2))
ggplot(cius_data, aes(x=gender))+
  geom_bar(stat = "count")
ggplot(cius_data, aes(x=official_language_proficient))+
  geom_bar(stat = "count")
ggplot(cius_data, aes(x=employment_status))+
  geom_bar(stat = "count")
ggplot(cius_data, aes(x=is_immigrant))+
  geom_bar(stat = "count")
```

graph of internet use status
```{r}
ggplot(cius_data, aes(x=uses_internet))+
  geom_bar(stat = "count")
```
frequency plot of hours used
```{r}
ggplot(cius_data, aes(x=hours_used))+
  geom_bar(stat = "count")
```
graph of access at home
```{r}
ggplot(cius_data, aes(x=home_internet_access))+
  geom_bar(stat = "count")
```
summary statistics of household connection type
```{r}
connection_data <- summarise_all(cius_data[,62:73], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "connection_type", values_to = "count")
ggplot(connection_data, aes(x=reorder(connection_type, -count),y=count))+
  geom_bar(stat = "identity")
```

frequency plot of connection speed

```{r}
ggplot(cius_data, aes(x=fct_infreq(connection_speed)))+
  geom_bar(stat = "count")
```

summary statistics of device type
```{r}
device_data <- summarise_all(cius_data[,15:21], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "device_type", values_to = "count")
ggplot(device_data, aes(x=reorder(device_type, -count),y=count))+
  geom_bar(stat = "identity")
```
summary statistics of communications
```{r}
communications_data <- summarise_all(cius_data[,22:28], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "communications_type", values_to = "count")
ggplot(communications_data, aes(x=reorder(communications_type, -count),y=count))+
  geom_bar(stat = "identity")
```
summary statistics of information
```{r}
information_data <- summarise_all(cius_data[,30:33], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "information_type", values_to = "count")
ggplot(information_data, aes(x=reorder(information_type, -count),y=count))+
  geom_bar(stat = "identity")
```
summary statistics of entertainment
```{r}
entertainment_data <- summarise_all(cius_data[,35:43], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "entertainment_type", values_to = "count")
ggplot(entertainment_data, aes(x=reorder(entertainment_type, -count),y=count))+
  geom_bar(stat = "identity")
```
summary statistics of commerce
```{r}
commerce_data <- summarise_all(cius_data[,45:48], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "commerce_type", values_to = "count")
ggplot(commerce_data, aes(x=reorder(commerce_type, -count),y=count))+
  geom_bar(stat = "identity")
```
summary statistics of other activities
```{r}
other_activities_data <- summarise_all(cius_data[,50:57], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "other_activities_type", values_to = "count")
ggplot(other_activities_data, aes(x=reorder(other_activities_type, -count),y=count))+
  geom_bar(stat = "identity")
```

graph of internet use as source of income

```{r}
ggplot(cius_data, aes(x=internet_income))+
  geom_bar(stat = "count")
```
graph of major or additional source of income

```{r}
income_data <- summarise_all(cius_data[,60:61], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "income_type", values_to = "count")
ggplot(income_data, aes(x=reorder(income_type, -count),y=count))+
  geom_bar(stat = "identity")
```

```{r}
internet_activity <- cbind(select(cius_data,
                                  no_communication,
                                  no_commerce,
                                  no_entertainment,
                                  no_information,
                                  no_other_activities,
                                  internet_income)) %>%
  summarise_all(table)
internet_activity <- internet_activity[-c(2:4),] %>%
  pivot_longer(everything(),names_to = "activity", values_to = "count")
ggplot(internet_activity, aes(x=reorder(activity, -count),y=count))+
  geom_bar(stat = "identity")
```