### 2022-04-27 outputs/paper/paper.Rmd
added results content

---
title: "Impact of Internet Infrastructure on Canadian Household Productivity and Recreation"
subtitle: "Brief Overview of Canadian Internet Use Trends"
author: "Haocong Wu"
date: "`r format(Sys.time(), '%d %B %Y')`"
thanks: "Code and data are available at: https://github.com/hcgw0318/2018-Canadian-Internet-Use-Analysis"
abstract: "Internet access is critical to many users. This paper examines the quality of internet access that users have and how they are impacted by subpar service. With data collected by Statistics Canada in the Canadian Internet Use Survey, and modelled with linear regression, age has a negative correlation with user household internet usage, whilst education level has a strong positive correlation."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(kableExtra)
library(bookdown)
library(here)
library(forcats)
library(modelsummary)
```

```{r, include=FALSE}
cius_data <- read_csv(here("outputs/data/cleaned_data_v2.csv"))
```

Keywords: internet use, internet infrastructure, statistics canada, canadian residents, linear regression

# Introduction

Internet access is becoming, if not already considered, a human right. It is a standard point of communication used by the majority of the Canadian population. The platform that the internet provides can be used to host a variety of activities, ranging from those of productive nature, to others of recreational and leisurely type. The growing importance of internet access necessitates a revision of the bare minimums of providing internet service.

This paper examines the causal relations between individual internet usage amounts to factors such as connection speed,nature of use, and necessity of use.

The following sections of the paper is organized as stated: the Data section discusses the source of the data and its surveying methodology, the Modelling section attempts to formalize a model incorporating various factors to explain the response variable of user household internet use duration, the Results section discusses the impact of the selected factors, and the Discussion section comments on the results and its validity, potential bias, and weaknesses of the paper.

# Data

## Data Source

The Canadian Internet Use Survey (CIUS) replaced the Household Internet Use Survey (HIUS) in 2005 [@citecius]. The 2018 of the CIUS aims to record and track the quickly increasing Canadian internet utilization. Aspects of internet technologies that Canadians interact with such as online government services, social networks, smartphone use, and online commercial platforms are presented to respondents for their sentiments.

Survey data was provided by the <odesi> portal, maintained by the Ontario Council of University Libraries.

The particular set of data used in this report is a subset of the topics covered by the 2018 CIUS.

## Survey Methodology

Based on the 2018 CIUS documentation [@citecius2018doc], the survey methodology can be summarized by the following.

The CIUS attempts to gauge the population habits regarding various facets of internet use The CIUS' stated target population includes all individuals at and above the age of 15 in Canada, excluding those in the Canadian Territories and excluding institutionalized individuals.

The sampling frame of the GSS relies on the availability of telephone access. Households without telephones, as well as households with telephone services not covered by the current frame, were excluded from the sampling frame.

The sampling process consisted of location based random sampling without replacement. A target minimum sample size for each province was determined, and this minimum sample size was allocated to the strata within the province.

Each respondent was contacted to complete either an electronic questionnaire, or participate in a computer assisted telephone interview. If the respondent refused or was absent, further contact attempts were made through telephone contact until reasonable confirmation of non-response, and the respondent is removed from the pool.

The 2018 CIUS had a target sample size of 15,000. The actual number of respondents was 13,810, the total households contacted was 33,248, and the overall response rate was 41.6%.

## Data Processing Tools and Methodology

Data processing and analysis in this project is done in the R statistical programming language [@citeR]. The following R packages were imported and used for their corresponding purpose:

* `tidyverse` [@citetidyverse] : Used for general logical commands, file importing, and data manipulation

* `knitr` [@citeknitr] : Used to generate and knit the Markdown document to PDF

* `ggplot2` [@citeggplot2] : Used to generate graphs and apply styling

* `kableExtra` [@citekableExtra] : Used to assist in table formatting

* `bookdown` [@citebookdown] : Used to improve the Markdown document formatting

* `here` [@citehere] : Used to simplify file access in a project environment

* `forcats` [@citeforcats] : Used to sort data tables by frequency

* `modelsummary` [@citemodelsummary] : Used to display regression model information

## Relevant variables

Variables such as whether the respondent used the internet and had an internet connection at home were used to identify the subset of the sample whose internet usage time is the response of interest.
Variables likely to explain variations in internet usage time contained in the survey were age, education level, and purpose of internet usage.

```{r, echo=FALSE}
demo_dat <- cius_data %>% select(age_group, gender) %>%
  group_by(age_group, gender) %>%
  tally() %>%
  pivot_wider(names_from = gender, values_from = n) %>% 
  remove_rownames %>% 
  column_to_rownames(var="age_group")

prov_dat <- cius_data %>% select(province) %>% group_by(province) %>% tally()
colnames(prov_dat) <- c("Province", "Count")
```

```{r demotab1, fig.cap="Survey Respondent Demographics - Location", echo=FALSE}
# table
kbl(prov_dat, caption = "Location Demographics", booktabs = T, linesep = "") %>%
  kable_styling(bootstrap_options = c("hover", "condensed"), latex_options = "HOLD_position") %>%
  kable_paper("striped", full_width = F)
```

```{r demotab2, fig.cap="Survey Respondent Demographics - Age and Gender", echo=FALSE}
# table
kbl(demo_dat, caption = "Age and Gender Demographics", booktabs = T, linesep = "") %>%
  kable_styling(bootstrap_options = c("hover", "condensed"), latex_options = "HOLD_position") %>%
  kable_paper("striped", full_width = F)
```

In Table \@ref(tab:demotab1) and Table \@ref(tab:demotab2), the survey demographics are shown. It can be seen that the sample population is skewed towards more female respondents, and skewed towards older respondents.

```{r, echo=FALSE}
connection_data <- summarise_all(cius_data[,62:73], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "connection_type", values_to = "count")
```

```{r homeconnection, fig.cap="Respondent home connection type", echo=FALSE}
ggplot(connection_data, aes(x=reorder(connection_type, -count),y=count))+
  geom_bar(stat = "identity") + 
  scale_y_continuous() +
  theme(axis.text.x = element_text(angle = -15, vjust = 1, hjust=0)) + 
  labs(x = "Connection type", 
       y = "Count",
       title = "Home internet connection type")
```

In Figure \@ref(fig:homeconnection), the three most popular types of home internet connections are cable, fiber optics, and mobile data. After that, approximately 7% of respondents are not sure what type of connection they have.

```{r, echo=FALSE}
internet_activity <- cbind(select(cius_data,
                                  no_communication,
                                  no_commerce,
                                  no_entertainment,
                                  no_information,
                                  no_other_activities,
                                  internet_income)) %>%
  summarise_all(table)
internet_activity <- internet_activity[-c(2:4),] %>%
  rename(Communication=no_communication,
         Commerce=no_commerce,
         Entertainment=no_entertainment,
         Information=no_information,
         Other=no_other_activities,
         Income=internet_income) %>%
  pivot_longer(everything(),names_to = "activity", values_to = "count")
```

```{r usepurpose, fig.cap="Respondent internet use purpose", echo=FALSE}
ggplot(internet_activity, aes(x=reorder(activity, -count),y=count))+
  geom_bar(stat = "identity") + 
  labs(x = "Internet usage type", 
       y = "Count",
       title = "Purposes of internet usage")
```

In Figure \@ref(fig:usepurpose), common uses of the internet are displayed.

```{r, echo=FALSE}
cius_data$hours_used <- cius_data$hours_used %>% 
  factor(levels=c("Less than 5 hours per week",
                  "5 to less than 10 hours per week",
                  "10 to less than 20 hours per week",
                  "20 to less than 40 hours per week",
                  "40 hours or more per week",
                  "Not stated",
                  "Valid skip"))
```

```{r internethours, fig.cap="Respondent weekly internet use", echo=FALSE}
ggplot(cius_data, aes(x=hours_used))+
  geom_bar(stat = "count") + 
  theme(axis.text.x = element_text(angle = -15, vjust = 1, hjust=0)) + 
  labs(x = "Duration of use",
       y = "Count",
       title = "Hours of internet use per week (excluding entertainment)")
```

In Figure \@ref(fig:internethours), the distribution of weekly internet usage of the respondents is shown. The most common amount of time spent on the internet per week is between 5 to 10 hours.

# Results

```{r internetuse, fig.cap="Respondent internet use and home internet access", echo=FALSE}
#graph
ggplot(cius_data, aes(x=fct_infreq(uses_internet), fill=home_internet_access))+
  geom_bar(stat = "count") +
  labs(x = "Uses the internet", 
       y = "Count",
       title = "Respondents and their usage and access to internet at home",
       fill = "Home internet access")
```

In Figure \@ref(fig:internetuse), it is shown that most users use the internet in some way. The majority of respondents that do not use the internet also do not have an internet connection at home.

```{r, echo=FALSE}
cius_data$education_level <- cius_data$education_level %>% 
  factor(levels=c("High school diploma or an equivalency certificate or less",
                  "Certificate/diploma Univ. below bachelor/College/CEGEP/trade",
                  "University degree above the bachelor or bachelor's degree",
                  "Not stated"))

```

```{r educationlevel, fig.cap="Respondent education level and internet use", echo=FALSE}
ggplot(cius_data, aes(x=education_level,fill=uses_internet))+
  geom_bar(stat = "count")+
  theme(axis.text.x = element_text(angle = -15, vjust = 1, hjust=0))+
  labs(x = "Education level", 
       y = "Count",
       title = "Education level and internet use",
       fill = "Internet usage")
```

In Figure \@ref(fig:educationlevel), it can be seen that internet usage proportion increases with respondent education level.

```{r, echo=FALSE}
cius_data$age_group <- cius_data$age_group %>% 
  factor(levels=c("15 to 24 years",
                  "25 to 34 years",
                  "35 to 44 years",
                  "45 to 54 years",
                  "55 to 64 years",
                  "65 years and over"))
```

```{r agegroup, fig.cap="Respondent age group and internet use", echo=FALSE}
ggplot(cius_data, aes(x=age_group,fill=uses_internet))+
  geom_bar(stat = "count") + 
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + 
  labs(x = "Age group", 
       y = "Count",
       title = "Age group and internet use",
       fill = "Internet usage")
```

In Figure \@ref(fig:agegroup), as age group increases, the proportion of the group that does not use the internet increases.

```{r, echo=FALSE, warning=FALSE}
set.seed(318)
regression_dat <- cius_data[cius_data$uses_internet=="Yes" &
                              cius_data$home_internet_access=="Yes" &
                              !(cius_data$hours_used=="Valid skip" | cius_data$hours_used=="Not stated"),]
regression_dat <- regression_dat %>% 
  mutate(hours_used=case_when(hours_used=="Less than 5 hours per week"~
                                sample(x = c(0:4), size = n(),replace = TRUE),
                              hours_used=="5 to less than 10 hours per week"~
                                sample(x = c(5:9), size = n(),replace = TRUE),
                              hours_used=="10 to less than 20 hours per week"~
                                sample(x = c(10:19), size = n(),replace = TRUE),
                              hours_used=="20 to less than 40 hours per week"~
                                sample(x = c(20:39), size = n(),replace = TRUE),
                              hours_used=="40 hours or more per week"~
                                sample(x = c(40:60), size = n(),replace = TRUE)))
regression_dat <- regression_dat %>%
  mutate_at(c("no_communication",
              "no_commerce",
              "no_entertainment",
              "no_information",
              "no_other_activities",
              "internet_income"), funs(case_when(.=="Yes"~TRUE,
                                                 TRUE~FALSE)))
regression_dat <- regression_dat %>%
  mutate_at(vars(62:73), funs(case_when(.=="Yes"~TRUE,
                                        TRUE~FALSE)))
regression_dat$connection_type <- names(regression_dat)[62:73][max.col(regression_dat[62:73], 'first')]

lm_model <- lm(data = regression_dat, hours_used ~ 
                 gender + 
                 education_level + 
                 age_group +
                 no_communication +
                 no_commerce +
                 no_entertainment +
                 no_information +
                 no_other_activities +
                 internet_income)
```

```{r regressionmodel, fig.cap="2018 Canadian Internet Use Regression Model", echo=FALSE}
modelsummary(lm_model, fmt=1, title = "Model of weekly internet use quantity")
```

In Table \@ref(tab:regressionmodel), the regression model of internet use hours as a function of gender, education level, age group, and internet activities is displayed.

# Discussion

## Greater proportion of female respondents

## Greater proportion of older respondents

## Education and age trends

## Weaknesses and improvements

# References